---
title: "Data Loading"
output:
  html_document:
    df_print: paged
---

# **01 - Data Loading and Cleaning**

------------------------------------------------------------------------

## Setup the Environment

```{r}

# Import packages reproducibily

pkgs <- c('tidyverse', 'readr', 'dplyr', 'lubridate')

for (p in pkgs) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p)
    library(p, character.only = TRUE)
  }
}

```

## Import Data from Device

Here, we load data from all three of our main datasets on the GBR into dataframe for analysis.

The datasets themselves will actually be local to our computers to avoid issues with pushing really large files; our folder structure will be mirrored so that the code works for each of us.

Brody manages a OneDrive folder which contains the full datasets so we both have the same versions in our personal "Data" folder.

```{r}

# Import all eReef Variables

CHEM_eReef <- read_csv("Data/eReef/eReef_CHEM_v2.csv")
PHYS_eReef <- read_csv("Data/eReef/eReef_PHYS_v2.csv")
  
# Import Fish Data

Fish_AIMS_manta <- read_csv("Data/Fish/manta-tow-by-reef.csv")  # Large AIMS dataset

# Specific AIMS fish Data

Agincourt_fish_model <- read_csv("Data/Fish/Agincourt_fish_modelled.csv") |>
  mutate(Date = date_decimal(date)) |>
  select(Date, mean)


Broomfield_fish_model <- read_csv("Data/Fish/Broomfield_fish_modelled.csv") |>
  mutate(Date = date_decimal(date)) |>
  select(Date, mean)

Rib_fish_model <- read_csv("Data/Fish/Rib_fish_modelled.csv") |>
  mutate(Date = date_decimal(date)) |>
  select(Date, mean)

```

**Complete Pivots, Joins, and Other Filtering**

```{r}

# Extract pH data

pH_all_sites <- CHEM_eReef |>
  filter(Variable == "PH") |>
  pivot_wider(names_from = Variable, values_from = mean) |>
  select(`Site Name`, `Aggregated Date/Time`, PH)

# Extract CO2 data

CO2_all_sites <- CHEM_eReef |>
  filter(Variable == "DIC") |>
  pivot_wider(names_from = Variable, values_from = mean) |>
  select(`Site Name`, `Aggregated Date/Time`, DIC)

# Split Chem Data

Chem_Agincourt <- CHEM_eReef |>
  filter(`Site Name` == "Agincourt Reef No 1")
Chem_Rib <- CHEM_eReef |>
  filter(`Site Name` == "Rib Reef")
Chem_Broomfield <- CHEM_eReef |>
  filter(`Site Name` == "Broomfield Reef")

# Split Phys Data

Phys_Agincourt <- PHYS_eReef |>
  filter(`Site Name` == "Agincourt Reef No 1")
Phys_Rib <- PHYS_eReef |>
  filter(`Site Name` == "Rib Reef")
Phys_Broomfield <- PHYS_eReef |>
  filter(`Site Name` == "Broomfield Reef")

# Format Fish Data

Fish_AIMS_manta_clean <- Fish_AIMS_manta  |>
  filter(REEF_NAME == "AGINCOURT REEFS (NO 1)" | REEF_NAME == "BROOMFIELD REEF" | REEF_NAME == "RIB REEF") |>
  select(SAMPLE_DATE, REEF_NAME, TOTAL_COTS, TOTAL_TROUT, MEAN_LIVE_CORAL)

# Aggregate Chem and Phys for each Site

WQ_Agincourt <- Phys_Agincourt |>
  bind_rows(Chem_Agincourt) |>
  arrange(`Aggregated Date/Time`) |>
  select(`Aggregated Date/Time`, Variable, `Site Name`, mean) |>
  pivot_wider(
    names_from = Variable,
    values_from = mean
  ) |>
  select(-omega_ar, -alk)

WQ_Broomfield <- Phys_Broomfield |>
  bind_rows(Chem_Broomfield) |>
  arrange(`Aggregated Date/Time`) |>
  select(`Aggregated Date/Time`, Variable, `Site Name`, mean) |>
  pivot_wider(
    names_from = Variable,
    values_from = mean
  ) |>
  select(-omega_ar, -alk)

WQ_Rib <- Phys_Rib |>
  bind_rows(Chem_Rib) |>
  arrange(`Aggregated Date/Time`) |>
  select(`Aggregated Date/Time`, Variable, `Site Name`, mean) |>
  pivot_wider(
    names_from = Variable,
    values_from = mean
  ) |>
  select(-omega_ar, -alk)
  

```

**Export the Cleaned Data**

```{r}

vars <- c(
  "Agincourt_fish_model", "Broomfield_fish_model",
  "CO2_all_sites", "Fish_AIMS_manta_clean",
  "pH_all_sites", "Rib_fish_model", "WQ_Agincourt", "WQ_Broomfield", "WQ_Rib"
)

for (v in vars) {
  saveRDS(get(v), paste0("Data/Cleaned_Data/", v, ".rds"))
}

```
