---
title: "R Notebook"
output: html_notebook
---

**Import Packages**

```{r}

# Import packages reproducibily

pkgs <- c('tidyverse', 'readr', 'dplyr', 'lubridate')

for (p in pkgs) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p)
    library(p, character.only = TRUE)
  }
}

```

**Import Data from Device**

Here, we load data from all three of our main datasets on the GBR into dataframe for analysis.

The datasets themselves will actually be local to our computers to avoid issues with pushing really large files; our folder structure will be mirrored so that the code works for each of us.

I (Brody) will create a OneDrive folder which contains the full datasets so we both have the same versions in our personal "Data" folder.

This is also the place to rename columns and perform necessary joins.

```{r}

# Import Water Temperature Data

Temp_Davies <- read_csv("Data/Temperature/Davies_Temp.csv") |>
  select(time, raw_value)

Temp_Heron <- read_csv("Data/Temperature/Heron_Temp.csv") |>
  select(time, raw_value)

Temp_Pandora <- read_csv("Data/Temperature/Pandora_Temp.csv") |>
  select(time, cal_val)


# Import Chemistry Variables

Chem_eReef <- read_csv("Data/eReef/CHEM_Vars.csv") # Full Chem df

# Import Physical Variables

Phys_eReef <- read_csv("Data/eReef/PHYS_Vars.csv") # Full Phys df

# Import Fish data

Davies_fish <- read_csv("Data/Fish/Davies_fish.csv")

Pandora_fish <- read_csv("Data/Fish/Pandora_fish.csv")

```

**Complete Pivots, Joins, and Other Filtering**

```{r}

# Extract pH data

pH_all_sites <- Chem_eReef |>
  filter(Variable == "PH") |>
  pivot_wider(names_from = Variable, values_from = mean) |>
  select(`Site Name`, `Aggregated Date/Time`, PH)

# Extract CO2 data

CO2_all_sites <- Chem_eReef |>
  filter(Variable == "DIC") |>
  pivot_wider(names_from = Variable, values_from = mean) |>
  select(`Site Name`, `Aggregated Date/Time`, DIC)

# Split Chem Data

Chem_Davies <- Chem_eReef |>
  filter(`Site Name` == "Davies Reef")
Chem_Heron <- Chem_eReef |>
  filter(`Site Name` == "Heron Island")
Chem_Pandora <- Chem_eReef |>
  filter(`Site Name` == "Pandora Reef")

# Split Phys Data

Phys_Davies <- Phys_eReef |>
  filter(`Site Name` == "Davies Reef")
Phys_Heron <- Phys_eReef |>
  filter(`Site Name` == "Heron Island")
Phys_Pandora <- Phys_eReef |>
  filter(`Site Name` == "Pandora Reef")

# Format Fish Data

Pandora_fish_clean <- Pandora_fish |>
  mutate(Date = decimal_date(as.POSIXct(date))) |>
  select(Date, mean)

Davies_fish_clean <- Davies_fish |>
  mutate(Date = decimal_date(as.POSIXct(survey_date))) |>
  select(Date, tows, cots)

```
