---
title: "Visualization and Exploration"
output:
  html_document:
    df_print: paged
---

# **02 - Exploration and Visualization**

------------------------------------------------------------------------

## Setup the Environment

```{r}

# Import packages reproducibily

pkgs <- c('tidyverse', 'readr', 'dplyr', 'lubridate', 'purrr', 'ggcorrplot', 'patchwork')

for (p in pkgs) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p)
    library(p, character.only = TRUE)
  }
}

remove(pkgs, p)

```

```{r}

# Load all RDS files from Cleaned_Data folder

files <- list.files("Data/Cleaned_Data", pattern = "\\.rds$", full.names = TRUE)

for (f in files) {
  name <- tools::file_path_sans_ext(basename(f))
  assign(name, readRDS(f))
}

```

## **Visualize the Cleaned Data to Identify Trends**

#### Explore fish population data

```{r}

# Combine the fish data

fish_combined <- bind_rows(
  Agincourt_fish_model %>% mutate(Site = "Agincourt"),
  Broomfield_fish_model %>% mutate(Site = "Broomfield"),
  Rib_fish_model %>% mutate(Site = "Rib")
)

# Plot all fish on the same plot

ggplot(fish_combined, aes(x = Date, y = mean, color = Site)) +
  geom_line() +
  labs(
    title = "Fish Population at Each Site Over Time",
    y = "Population",
    x = "Date"
  ) +
  theme_minimal()


```

### Monthly Matrices

```{r}
# Combine water quality datasets
WQ_all <- WQ_Agincourt |>
  bind_rows(WQ_Broomfield, WQ_Rib)

# Filter fish dataset to align with water quality set
fish_combined_filter <- fish_combined |>
  mutate(Date = as.Date(Date)) |>
  filter(Date > as.Date("2010-12-01"),
         Date < as.Date("2019-12-31")) |>
  rename(Fish = mean)

# Rename
WQ_all <- WQ_all |>
  mutate(Date = as.Date(`Aggregated Date/Time`)) |>
  rename(Site = `Site Name`) |>
  mutate(
    Site = case_when(
      Site == "Agincourt Reef No 1" ~ "Agincourt",
      Site == "Broomfield Reef" ~ "Broomfield",
      Site == "Rib Reef" ~ "Rib",
      TRUE ~ Site
    )
  )

```

```{r}
# Define function to generate correlation matrices for variables average over prior
# number of months (m)
monthly_matrices <- function(fish_data, WQ_all, max_month = 24) {
  
  map_dfr(1:max_month, function(m) {
    
    fish_WQ_means <- fish_data |>
      rowwise() |>
      mutate(
        Temp_avg = {
          d <- Date; s <- Site
          mean(
            WQ_all |>
              filter(Site == s, Date < d, Date >= d %m-% months(m)) |>
              pull(temp),
            na.rm = TRUE
          )
        },
        Salt_avg = {
          d <- Date; s <- Site
          mean(
            WQ_all |>
              filter(Site == s, Date < d, Date >= d %m-% months(m)) |>
              pull(salt),
            na.rm = TRUE
          )
        },
        PH_avg = {
          d <- Date; s <- Site
          mean(
            WQ_all |>
              filter(Site == s, Date < d, Date >= d %m-% months(m)) |>
              pull(PH),
            na.rm = TRUE
          )
        },
        DIC_avg = {
          d <- Date; s <- Site
          mean(
            WQ_all |>
              filter(Site == s, Date < d, Date >= d %m-% months(m)) |>
              pull(DIC),
            na.rm = TRUE
          )
        }
      ) |>
      ungroup()
    
    cor_matrix <- fish_WQ_means |>
      select(where(is.numeric)) |>
      drop_na() |>
      cor()
    
    as.data.frame(as.table(cor_matrix)) |>
      rename(r = Freq) |>
      mutate(month_window = m)
  })
}

```

```{r}
# Run function to create output dataframe
cor_results <- monthly_matrices(
  fish_data = fish_combined_filter,
  WQ_all = WQ_all,
  max_month = 24) |>
  filter(Var1 != Var2)

```

```{r}
# Create dataframe of strongest correlations
strongest_pairs <- cor_results |>
  
  # 1. remove self-correlations
  filter(Var1 != Var2) |>
  
  # 2. create ordered pair to remove duplicates
  mutate(
    var_low  = pmin(as.character(Var1), as.character(Var2)),
    var_high = pmax(as.character(Var1), as.character(Var2))
  ) |>

  # 3. group by unique variable pair
  group_by(var_low, var_high) |>
  
  # 4. find strongest magnitude across months
  mutate(abs_r = abs(r)) |>
  slice_max(abs_r, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(
    Var1 = var_low,
    Var2 = var_high,
    r,
    month_window
  )

strongest_pairs

```

```{r}
ggplot(strongest_pairs, 
       aes(x = Var1, y = Var2, fill = r)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0
  ) +
  theme_minimal() +
  labs(
    title = "Strongest Correlation (Across 1–24 Month Windows)",
    fill = "Correlation (r)"
  ) +
  coord_equal()
```

### Plot each relationship over optimal averaging window:

```{r}
compute_means <- function(fish_data, WQ_all, m) {
  fish_data |>
    rowwise() |>
    mutate(
      Temp_avg = {
        d <- Date; s <- Site
        mean(WQ_all |>
               filter(Site == s,
                      Date < d,
                      Date >= d %m-% months(m)) |>
               pull(temp),
             na.rm = TRUE)
      },
      Salt_avg = {
        d <- Date; s <- Site
        mean(WQ_all |>
               filter(Site == s,
                      Date < d,
                      Date >= d %m-% months(m)) |>
               pull(salt),
             na.rm = TRUE)
      },
      PH_avg = {
        d <- Date; s <- Site
        mean(WQ_all |>
               filter(Site == s,
                      Date < d,
                      Date >= d %m-% months(m)) |>
               pull(PH),
             na.rm = TRUE)
      },
      DIC_avg = {
        d <- Date; s <- Site
        mean(WQ_all |>
               filter(Site == s,
                      Date < d,
                      Date >= d %m-% months(m)) |>
               pull(DIC),
             na.rm = TRUE)
      }
    ) |>
    ungroup()
}

```

```{r}
data_averages_all <- pmap_dfr(
  strongest_pairs,
  function(Var1, Var2, r, month_window) {
    
    df <- compute_means(fish_combined_filter, WQ_all, month_window)
    
    tibble(
      x = df[[Var1]],
      y = df[[Var2]],
      Var1 = Var1,
      Var2 = Var2,
      r = r,
      month_window = month_window
    )
  }
)
```

```{r, fig.width=14, fig.height=18}
ggplot(data_averages_all, aes(x = x, y = y)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~ Var1 + Var2, ncol = 2, scales = "free") +
  theme_minimal()
```

#### Summary of Current Analysis Performed

The first figure is a simple exploratory look at the fish abundance datasets. The data is plotted against time for the entire duration of the sample, with colors denoting the different research sites. Clear abundance spikes appear in the years after 2020, suggesting some sort of COVID-19 related fishing changes.

The second figure displays the correlation between different chemical variables and fish abundance, over both 3 and 6 month rolling average periods. The results show a stronger correlation between fish abundance and PH over a longer rolling average period, suggesting that pH changes may create their biggest effects over longer timescales.

While not shown in the figures, we have also examined much of the raw data to determine what timescale would be best for this analysis. Sometime between 2010 - 2023 seems to have the most quality data.

### Combine fish abundance with chemical and physical variables

```{r}

# Summarize chem to yearly means

chem_yearly <- WQ_all |>
  mutate(Year = year(Date)) |>
  group_by(Site, Year) |>
  summarise(across(c(temp, salt, PH, DIC), mean, na.rm = TRUE), .groups = "drop")

# Summarize the fish data to unified yearly data

fish_combined_yearly <- fish_combined |>
  mutate(Year = year(Date))

# Join to fish data by Site and Year

All_fish_all_vars <- fish_combined_yearly |>
  left_join(chem_yearly, by = c("Site", "Year"))

```

------------------------------------------------------------------------

## [Final analysis steps:]{.underline}

### Connect pH to fish abundance

Futher determine the correlation between fish abundance and pH via more correlation matrices:

```{r}



```

Create various regression models to examine correlation coefficients across the variables:

```{r}



```

### Connect DIC to water pH

Run more correlation analysis on the DIC and pH:

```{r, fig.width=4, fig.height=4}

DIC_pH <- data_averages_all |>
  filter(Var1 == "DIC_avg", Var2 == "PH_avg")

pH_Fish <- data_averages_all |>
  filter(Var1 == "Fish", Var2 == "PH_avg")

ggplot(DIC_pH, aes(x = x, y = y)) +
  geom_point(color = "mediumaquamarine", size = 2) +
  geom_smooth(method = "lm", col = "mediumaquamarine") +
  labs(
    title = "DIC and pH across Sites and Time",
    subtitle = paste("r =", round(unique(DIC_pH$r), 2)),
    x = "Dissolved Inorganic Carbon",
    y = "pH"
  ) +
  theme_minimal()

ggplot(pH_Fish, aes(x = y, y = x)) +
  geom_point(color = "coral", size = 2) +
  geom_smooth(method = "lm", col = "coral") +
  labs(
    title = "pH and Fish across Sites and Time",
    subtitle = paste("r =", round(unique(pH_Fish$r), 2)),
    x = "pH",
    y = "Fish Abundance"
  ) +
  theme_minimal()

```

```{r, fig.width=5, fig.height=2}
fish_round <- fish_combined |>
  mutate(round = round(mean, 5))

pH_fish_ts <- pH_Fish |>
  select(-Var1, -Var2) |>
  rename(Fish = x, PH_avg = y) |>
  mutate(Fish = round(Fish, 5)) |>
  left_join(fish_round, by = c("Fish" = "round"))

ggplot(pH_fish_ts, aes(x = Date, y = Fish, col = Site)) +
  geom_line() +
  theme_minimal() +
  labs(
    x = "Date",
    y = "Fish Abundance",
    color = "Site") + 
  theme(legend.position = "none")


```

```{r, fig.width=5, fig.height=2.5}
ggplot(pH_fish_ts, aes(x = Date, y = PH_avg, col = Site)) +
  geom_line() +
  theme_minimal() +
  labs(
    x = "Date",
    y = "pH",
    color = "Site") + 
  theme(legend.position = "bottom")
```

### Conclude that DIC correlates with fish abundance

Therefore, fish abundance is directly affected by rising atmospheric CO2:

```{r}



```
